### 2020-09-14 exploratory/2020/sub3_fit_subsets.Rmd
Clean up x axis labels

---
title: "Exploratory Plots"
date: "Sept. 2020"
output: 
  github_document:
    html_preview: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      fig.path = "figures/"
                      )
```
```{r lib}
library(cowplot)
library(ggtext)
library(here)
library(knitr)
library(tidyverse)
theme_set(theme_classic())
color_palette <- RColorBrewer::brewer.pal(4, "Dark2")
dataset_colors <- c(
  human = color_palette[[3]],
  marine = color_palette[[1]],
  mouse = color_palette[[4]],
  soil = color_palette[[2]]
)
```
```{r}
sensspec_fit <-
  read_tsv(here('subworkflows/3_fit_sample_subset/results/sensspec.tsv'))
```

## Subsetting datasets with different reference sizes

```{r fit_ref_frac}
sensspec_fit %>%
  ggplot(aes(x = as.factor(ref_frac), y = mcc, color = dataset)) +
  geom_boxplot() +
  scale_color_manual(values = dataset_colors) +
  ylim(0.5, 1) +
  facet_grid(method ~ ref_weight) +
  labs(title='OptiFit Performance',
       x='reference fraction')
```

sample fraction = 0.2

## Runtime
```{r runtime_ref_seqs}
runtime <- full_join(
  read_tsv(here('subworkflows/3_fit_sample_subset/results/benchmarks.tsv')),
  read_tsv(here('subworkflows/3_fit_sample_subset/results/input_sizes.tsv'))
  ) %>% 
  full_join(read_tsv(here('subworkflows/3_fit_sample_subset/results/gap_counts.tsv'))) %>% 
  mutate(num_total_seqs = num_ref_seqs + num_sample_seqs,
         gaps_frac = n_gaps / total_chars)
runtime %>% ggplot(aes(x=num_ref_seqs, y=s, color=dataset)) +
  geom_point(alpha = 0.3) +
  facet_grid(method ~ ref_weight) +
  scale_color_manual(values = dataset_colors) +
  scale_x_continuous(breaks = seq(0, 200000, 50000),
                     labels = c('0', '50k', '100k', '150k', '200k')) +
  labs(title = 'Runtime over reference size',
       x = '# sequences in reference',
       y = 'seconds')
```

```{r runtime_numotus}
runtime %>% full_join(sensspec_fit) %>% 
  ggplot(aes(x=numotus, y=s, color=dataset, shape = ref_weight)) +
  geom_point(alpha = 0.5) +
  facet_grid(method ~ ref_weight) +
  scale_color_manual(values = dataset_colors) +
  scale_x_continuous(breaks = seq(0, 40000, 10000),
                     labels = c('0', '10k', '20k', '30k', '40k')) +
  labs(title = 'Runtime over number of OTUs',
       x = '# OTUs',
       y = 'seconds')
```

## Fraction of sequences that map to the reference
```{r fraction_reads_mapped}
fractions <- read_tsv(here('subworkflows/3_fit_sample_subset/results/fraction_reads_mapped.tsv'))
fractions %>% 
  group_by(dataset, ref_weight, ref_frac) %>% 
  ggplot(aes(x=ref_frac, y=fraction_mapped, color=dataset)) +
  geom_jitter(alpha = 0.5, width = 0.01, size=1) +
  scale_color_manual(values = dataset_colors) +
  facet_wrap("ref_weight") +
  ylim(0, 1) +
  labs(title="Sequences mapped during closed-reference OptiFit",
       x='reference fraction',
       y='fraction mapped')
```