### 2020-10-05 exploratory/2020/sub3_fit_subsets_fixed.Rmd
Plot de novo results alongside optifit

---
title: "Exploratory Plots"
date: "Sept. 2020"
output: 
  github_document:
    html_preview: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      fig.path = "figures/"
                      )
```
```{r lib}
library(cowplot)
library(ggtext)
library(here)
library(knitr)
library(tidyverse)
theme_set(theme_classic())
color_palette <- RColorBrewer::brewer.pal(4, "Dark2")
dataset_colors <- c(
  human = color_palette[[3]],
  marine = color_palette[[1]],
  mouse = color_palette[[4]],
  soil = color_palette[[2]]
)
full_join_files <- function(filenames) {
  filenames %>% 
  lapply(read_tsv) %>% 
  reduce(full_join)
}
```
```{r download_data}
dat <- c('https://raw.githubusercontent.com/SchlossLab/OptiFitAnalysis/4fa7127bb6bacaa63657f0c18c0a8c6b6d18ebe7/subworkflows/3_fit_sample_subset/results/sensspec.tsv?token=AEHR6TOTYRF3XTCACDHGFRK7QRYH2',
              'https://raw.githubusercontent.com/SchlossLab/OptiFitAnalysis/4fa7127bb6bacaa63657f0c18c0a8c6b6d18ebe7/subworkflows/3_fit_sample_subset/results/benchmarks.tsv?token=AEHR6TLJWOKI4YC6OJ6BRMC7QRYXY',
              'https://raw.githubusercontent.com/SchlossLab/OptiFitAnalysis/4fa7127bb6bacaa63657f0c18c0a8c6b6d18ebe7/subworkflows/3_fit_sample_subset/results/input_sizes.tsv?token=AEHR6TJVM5FBTZRRY4USOYS7QRY6U',
              'https://raw.githubusercontent.com/SchlossLab/OptiFitAnalysis/4fa7127bb6bacaa63657f0c18c0a8c6b6d18ebe7/subworkflows/3_fit_sample_subset/results/gap_counts.tsv?token=AEHR6TMQKLHG2RNGUZQI35C7QRY5G',
              'https://raw.githubusercontent.com/SchlossLab/OptiFitAnalysis/4fa7127bb6bacaa63657f0c18c0a8c6b6d18ebe7/subworkflows/3_fit_sample_subset/results/fraction_reads_mapped.tsv?token=AEHR6TIY56KPVNGRKCOAYI27QRZBW') %>% 
  full_join_files() %>% 
  mutate(num_total_seqs = num_ref_seqs + num_sample_seqs,
         gaps_frac = n_gaps / total_chars)

denovo <- c('https://raw.githubusercontent.com/SchlossLab/OptiFitAnalysis/4fa7127bb6bacaa63657f0c18c0a8c6b6d18ebe7/subworkflows/1_prep_samples/results/sensspec.tsv?token=AEHR6TOXMAB75QDAO7CGTA27QR4KQ',
            'https://raw.githubusercontent.com/SchlossLab/OptiFitAnalysis/4fa7127bb6bacaa63657f0c18c0a8c6b6d18ebe7/subworkflows/1_prep_samples/results/benchmarks.tsv?token=AEHR6TI35D2TNWARGGL37W27QR4MM',
            'https://raw.githubusercontent.com/SchlossLab/OptiFitAnalysis/4fa7127bb6bacaa63657f0c18c0a8c6b6d18ebe7/subworkflows/1_prep_samples/results/dataset_sizes.tsv?token=AEHR6TJKE2C7SRDUZSGRTCC7QR4MG') %>% 
  full_join_files() %>% 
  mutate(ref_frac = 1)

denovo_sum <- denovo %>% group_by(dataset) %>% summarize(mean_mcc=mean(mcc), 
                                                         sd_mcc=sd(mcc),
                                                         mean_sec=mean(s),
                                                         sd_sec=sd(s),
                                                         num_ref_seqs = mean(num_seqs))
denovo_sum
```

## Subsetting datasets with different reference sizes

Sample fraction of 0.2 with 20 different seeds.

```{r fit_ref_frac}
dat %>%
  ggplot(aes(x = ref_frac, y = mcc, color = dataset)) +
  geom_jitter(size = 1, alpha = 0.3, width = 0.01) +
  stat_summary(fun = mean, geom = 'crossbar') +
  scale_color_manual(values = dataset_colors) +
  facet_grid(method ~ ref_weight) +
  labs(title='OptiFit Performance',
       x='reference fraction')
dat %>%
  ggplot(aes(x = ref_frac, y = mcc, color = dataset)) +
  geom_jitter(size = 1, alpha = 0.3, width = 0.01) +
  stat_summary(fun = mean, geom = 'crossbar') +
  geom_hline(aes(yintercept = mean_mcc, color = dataset), denovo_sum) +
  scale_color_manual(values = dataset_colors) +
  facet_grid(method ~ ref_weight) +
  labs(title='OptiFit Performance',
       x='reference fraction',
       caption='Sample fraction = 0.2 for OptiFit runs. The horizontal line represents _de novo_ clustering for the whole dataset.') +
  theme(plot.caption = element_markdown())
dat %>%
  ggplot(aes(x = ref_frac, y = mcc, color = ref_weight)) +
  geom_jitter(size = 1, alpha = 0.3, width = 0.01) +
  stat_summary(fun = mean, geom = 'crossbar') +
  geom_hline(aes(yintercept = mean_mcc), denovo_sum) +
  facet_grid(dataset ~ method) +
  labs(title='OptiFit Performance',
       x='reference fraction',
       caption='Sample fraction = 0.2 for OptiFit runs. The horizontal line represents _de novo_ clustering for the whole dataset.') +
  theme(plot.caption = element_markdown())
```

The performance is very consistent across reference fractions when weighting sequences by abundance or not at all (simple).
When weighting by distance, the performance decreases as the reference fraction increases.
This is expected because a larger fraction means **more** sequences with **fewer** 
pairwise distances under the 0.03 threshold are included in the reference.

## Runtime

```{r runtime_ref_seqs}
dat %>% ggplot(aes(x=num_ref_seqs, y=s, color=dataset)) +
  geom_point(alpha = 0.3) +
  facet_grid(method ~ ref_weight) +
  scale_color_manual(values = dataset_colors) +
  scale_x_continuous(breaks = seq(0, 200000, 50000),
                     labels = c('0', '50k', '100k', '150k', '200k')) +
  labs(title = 'Runtime over reference size',
       x = '# sequences in reference',
       y = 'seconds')
dat %>% group_by(num_ref_seqs, dataset, ref_weight, method) %>% 
  summarize(mean_sec=mean(s)) %>% 
  ggplot(aes(x=num_ref_seqs, y=mean_sec, color=ref_weight)) +
  geom_point(alpha = 0.5) +
  geom_point(aes(x=num_ref_seqs, y=mean_sec, color=NULL), denovo_sum) +
  facet_grid(dataset ~ method, scales = 'free') +
  #scale_color_manual(values = dataset_colors) +
  scale_x_continuous(breaks = seq(0, 200000, 50000),
                     labels = c('0', '50k', '100k', '150k', '200k')) +
  labs(title = 'Runtime over reference size',
       x = '# sequences in reference',
       y = 'seconds',
       caption = 'Sample fraction = 0.2 for OptiFit runs. The black dot represents _de novo_ clustering for the whole dataset') +
  theme(plot.caption = element_markdown())
```

Runtime is also fairly consistent regardless of weighting method.
For abundance and simple weighting, runtime seems to continue increasing roughly linearly (maybe a little worse than linear).
When weighting by distance, runtime reaches a plateau sooner than other selection methods as the size of the reference increases.
Perhaps the sequences with the most similarities are the most influential
over OTU clustering, 
so after you reach some critical number of similarities, the runtime doesn't increase as much.
I'm not sure why the human dataset takes so much longer than the others.

```{r runtime_numotus}
dat %>%
  ggplot(aes(x=numotus, y=s, color=dataset, shape = ref_weight)) +
  geom_point(alpha = 0.5) +
  facet_grid(method ~ ref_weight) +
  scale_color_manual(values = dataset_colors) +
  scale_x_continuous(breaks = seq(0, 40000, 10000),
                     labels = c('0', '10k', '20k', '30k', '40k')) +
  labs(title = 'Runtime over number of OTUs',
       x = '# OTUs',
       y = 'seconds')
```

Open-reference clustering has more OTUs because it will cluster any sequences 
_de novo_ that do not fit into existing OTUs, 
while closed-reference clustering throws them out.
This plot is confounded by the number of sequences in the reference, 
perhaps I should re-plot it as the ratio of OTUs to reference sequences?

## Fraction of sequences that map to the reference
```{r fraction_reads_mapped}
dat %>% 
  filter(method == 'closed') %>% 
  group_by(dataset, ref_weight, ref_frac) %>% 
  ggplot(aes(x=ref_frac, y=fraction_mapped, color=dataset)) +
  geom_jitter(alpha = 0.5, width = 0.01, size=1) +
  scale_color_manual(values = dataset_colors) +
  facet_wrap("ref_weight") +
  ylim(0, 1) +
  labs(title="Sequences mapped during closed-reference OptiFit",
       x='reference fraction',
       y='fraction mapped')
dat %>% 
  filter(method == 'closed') %>% 
  group_by(dataset, ref_weight, ref_frac) %>% 
  ggplot(aes(x=ref_frac, y=fraction_mapped, color=ref_weight)) +
  geom_jitter(alpha = 0.5, width = 0.01, size=1) +
  #scale_color_manual(values = dataset_colors) +
  facet_wrap("dataset", nrow=1) +
  ylim(0, 1) +
  labs(title="Sequences mapped during closed-reference OptiFit",
       x='reference fraction',
       y='fraction mapped')
```

The fraction of sequences that are able to be fit to the reference in closed-reference
clustering increases as the reference size increases when weighting sequences 
by abundance or not at all for reference selection.
It's also remarkably stable between abundance & simple weighting, and 
between all seeds.
Weighting by distance performs poorly, likely because the top n sequences with 
the most pairwise similarities are included in the reference, 
leaving sequences with few similarities to then be fit to the highly-connected reference.