### 2020-11-08 exploratory/2020-11/plots.Rmd

---
date: 'Nov. 2020'
output: 
  github_document:
    html_preview: false
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      fig.path = "figures/"
                      )
```
```{r lib}
library(cowplot)
library(ggtext)
library(glue)
library(here)
library(knitr)
library(tidyverse)
theme_set(theme_classic())
color_palette <- RColorBrewer::brewer.pal(4, "Dark2")
dataset_colors <- c(
  human = color_palette[[3]],
  marine = color_palette[[1]],
  mouse = color_palette[[4]],
  soil = color_palette[[2]]
)
set.seed(2018)
```

## Fit to reference databases

```{r}
ref_sizes <- read_tsv(here('subworkflows/2_fit_reference_db/results/ref_sizes.tsv'))
optifit_dbs <- read_tsv(here('subworkflows/2_fit_reference_db/results/optifit_dbs_results.tsv'))
```

```{r}
optifit_dbs
```