### 2020-11-11 exploratory/2020-11/plots.Rmd
Initial plots from updated results

---
date: 'Nov. 2020'
output: 
  github_document:
    html_preview: false
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      fig.path = "figures/"
                      )
```
```{r lib}
set.seed(2018)
library(cowplot)
library(ggtext)
library(glue)
library(here)
library(knitr)
library(tidyverse)

theme_set(theme_classic())
color_palette <- RColorBrewer::brewer.pal(4, "Dark2")
dataset_colors <- c(
  human = color_palette[[3]],
  marine = color_palette[[1]],
  mouse = color_palette[[4]],
  soil = color_palette[[2]]
)

mutate_perf <- function(dat) {
  dat %>% 
    mutate(mem_mb = max_rss,
           mem_gb = mem_mb / 1024) %>% 
    rename(sec = s,
           num_otus = sobs)
}
```

## _de novo_ clustering

```{r load_opticlust}
opticlust <- read_tsv(here('subworkflows/1_prep_samples/results/opticlust_results.tsv')) %>% 
  full_join(read_tsv(here('subworkflows/1_prep_samples/results/dataset_sizes.tsv'))) %>% 
  mutate_perf()
sum_opticlust <- opticlust %>% 
  group_by(dataset) %>% 
  summarize(mean_mcc = mean(mcc),
            sd_mcc = sd(mcc),
            mean_sec = mean(sec),
            sd_sec = sd(sec),
            mean_npshannon = mean(npshannon))
```
```{r mcc_opticlust}
opticlust %>% 
  ggplot(aes(dataset, mcc)) +
  geom_jitter(alpha = 0.5) +
  ylim(0.5, 1)
```
```{r runtime_opticlust}
opticlust %>% 
  ggplot(aes(num_seqs, sec, color = dataset)) +
  geom_point(alpha = 0.5)
opticlust %>% 
  ggplot(aes(num_otus, sec, color = dataset)) +
  geom_point(alpha = 0.5)
opticlust %>% 
  ggplot(aes(npshannon, sec, color = dataset)) +
  geom_point(alpha = 0.5)
```
## fit to reference databases

```{r load_fit-dbs}
ref_sizes <- read_tsv(here('subworkflows/2_fit_reference_db/results/ref_sizes.tsv'))
optifit_dbs <- read_tsv(here('subworkflows/2_fit_reference_db/results/optifit_dbs_results.tsv')) %>% 
  mutate_perf()
head(optifit_dbs)
```

```{r mcc_fit-db}
optifit_dbs %>% 
  ggplot(aes(x=method, y=mcc, color=ref)) +
  geom_hline(aes(yintercept = mean_mcc), sum_opticlust) +
  geom_jitter(alpha = 0.5) + 
  facet_wrap('dataset') +
  ylim(0, 1) +
  labs(title='OTU Quality: OptiFit with reference databases',
       caption='Black line: _de novo_ clustering') +
  theme(plot.caption = element_markdown())
```

```{r fraction-mapped_fit-db}
optifit_dbs %>% filter(method == 'closed') %>% 
  ggplot(aes(x=dataset, y=fraction_mapped, color=ref)) +
  geom_jitter(alpha = 0.5) + 
  ylim(0, 1) +
  labs(title="Sequences mapped during closed-reference OptiFit") +
  theme(plot.caption = element_markdown())
```
```{r runtime_fit-db}
optifit_dbs %>% 
  ggplot(aes(x=method, y=sec, color=ref)) +
  geom_jitter(alpha = 0.5) +
  geom_hline(aes(yintercept = mean_sec), sum_opticlust) +
  facet_wrap('dataset', scales = 'free') +
  labs(title='OptiFit runtime with reference databases',
       caption='Black line: _de novo_ clustering on the whole dataset') +
  theme(plot.caption = element_markdown())
```

## fit split datasets

```{r load_fit-split}
optifit_split <- read_tsv(here('subworkflows/3_fit_sample_split/results/optifit_split_results.tsv')) %>% 
  mutate_perf()
```

```{r mcc_fit-split}
optifit_split %>% 
  ggplot(aes(x = ref_frac, y = mcc, color = ref_weight)) +
  geom_jitter(size = 1, alpha = 0.3, width = 0.01) +
  stat_summary(fun = mean, geom = 'crossbar', alpha = 0.5) +
  geom_hline(aes(yintercept = mean_mcc), sum_opticlust) +
  facet_grid(dataset ~ method) +
  ylim(0, 1) +
  labs(title='OTU Quality: OptiFit with split datasets',
       x='reference fraction',
       caption='Black line: _de novo_ clustering on the whole dataset') +
  theme(plot.caption = element_markdown())
```

```{r runtime_fit-split}
optifit_split %>% 
  ggplot(aes(x = ref_frac, y = sec, color = ref_weight)) +
  geom_jitter(size = 1, alpha = 0.3, width = 0.01) +
  stat_summary(fun = mean, geom = 'crossbar', alpha = 0.5) +
  geom_hline(aes(yintercept = mean_sec), sum_opticlust) +
  facet_grid(dataset ~ method, scales = 'free') +
  labs(title='Runtime: OptiFit with split datasets',
       x='reference fraction',
       caption='Black line: _de novo_ clustering on the whole dataset') +
  theme(plot.caption = element_markdown())
```

```{r diversity_fit-split}
optifit_split %>% 
  ggplot(aes(x = ref_frac, y = npshannon, color = ref_weight)) +
  geom_jitter(size = 1, alpha = 0.3, width = 0.01) +
  stat_summary(fun = mean, geom = 'crossbar', alpha = 0.5) +
  geom_hline(aes(yintercept = mean_npshannon), sum_opticlust) +
  facet_grid(dataset ~ method, scales = 'free') +
  labs(title='Diversity: OptiFit with split datasets',
       x='reference fraction',
       caption='Black line: _de novo_ clustering on the whole dataset') +
  theme(plot.caption = element_markdown())
```

## vsearch

For reference-based clustering, datasets were fit to the greengenes database.

```{r load_vsearch}
vsearch <- read_tsv(here('subworkflows/4_vsearch/results/vsearch_results.tsv')) %>% 
  mutate_perf() %>% 
  full_join(opticlust) %>% 
  full_join(optifit_dbs %>% filter(ref == 'gg'))
```

```{r mcc_vsearch}
vsearch %>% 
  ggplot(aes(x = method, y = mcc, color = tool)) +
  geom_boxplot() +
  facet_wrap('dataset') +
  ylim(0, 1) + 
  labs(x = '')
```

```{r sec_vsearch}
vsearch %>% 
  ggplot(aes(x = method, y = sec, color = tool)) +
  geom_boxplot() +
  facet_wrap('dataset') +
  labs(x = '')
```

```{r diversity_vsearch}
vsearch %>% 
  ggplot(aes(x = method, y = npshannon, color = tool)) +
  geom_boxplot() +
  facet_wrap('dataset') +
  labs(x = '')
```

```{r fraction-mapped_vsearch}
vsearch %>% filter(method == 'closed') %>% 
  ggplot(aes(x = method, y = fraction_mapped, color = tool)) +
  geom_boxplot() + 
  facet_wrap('dataset') +
  ylim(0, 1) +
  labs(title="Sequences mapped during closed-reference OptiFit") +
  theme(plot.caption = element_markdown())
```