### 2020-12-10 paper/head.tex

\begin{center}
\vspace{25mm}

Kelly L. Sovacool${^1}$, Sarah L. Westcott${^2}$, M. Brodie Mumphrey${^1}$, Gabrielle A. Dotson${^1}$, Patrick D. Schloss${^2\dagger}$

$1$ Department of Computational Medicine and Bioinformatics,
University of Michigan, Ann Arbor, MI 48109

$2$ Department of Microbiology and Immunology,
University of Michigan, Ann Arbor, MI 48109

\vspace{30mm}

$\dagger$ To whom correspondence should be addressed: pschloss@umich.edu

\end{center}

\newpage
\linenumbers