### 2020-12-14 docs/slides/2020-12-14_lab-mtg.Rmd
Create slides for lab mtg pres.

---
title: "OptiFit project update"
subtitle: "+ a case study in debugging"
author: "Kelly Sovacool"
date: "2020-12-14 (updated: `r Sys.Date()`)"
output:
  xaringan::moon_reader:
    css: [default, default-fonts]
    lib_dir: libs
    nature:
      ratio: '16:9'
      highlightStyle: github
      highlightLines: true
      #titleSlideClass: ['center', 'middle', 'title-slide']
    chakra: libs/remark-latest.min.js
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(echo = FALSE,
                  message = FALSE
                  )
```

```{r chakra, eval = FALSE}
# run once to download remark-lastest
xaringan::summon_remark(to = 'docs/slides/libs')
```

```{r libs}
library(here)
library(knitr)
library(tidyverse)
xaringanExtra::use_share_again()
```

## OTU Clustering methods

- _De novo_ (OptiClust)
  - Cluster sequences based on their similarity.
  - OTU's aren't stable; you'll get different OTU assignments with different random seeds.
- Reference-based (OptiFit)
  - First cluster reference sequences _de novo_, then fit query sequences to those OTUs.

.footnote[
  [Westcott & Schloss (2015) _PeerJ_ ](https://doi.org/10.7717%2Fpeerj.1487);
  [Schloss (2016) _mSystems_ ](https://doi.org/10.1128%2FmSystems.00027-16);
  [Westcott & Schloss (2017) _mSphere_](10.1128/mSphereDirect.00073-17)
]

--

### Reference-based strategies

- Modes
  - Open - sequences that can't be fit to the reference are then clustered _de novo_ to create additional OTUs.
  - Closed - sequences that can't be fit to the reference are thrown out.
- Reference choice
  - External database (Silva, greengenes, RDP).
  - Split the dataset into a reference & query set.

---

# Changes since a couple months ago

- Sarah Westcott fixed some bugs in mothur & made OptiFit faster.
- I re-ran everything with silva v132.
- Ran vsearch on all the datasets
- Discovered a bug in my workflow code (last week!)

---

# Reference clustering against a database

.pull-left[
```{r mcc_fit-db}
include_graphics(here('exploratory', '2020-11_Nov-Dec', 'figures', 'mcc_fit-db-1.png'))
```
]

--

.pull-right[
```{r fraction-mapped_fit-db}
include_graphics(here('exploratory', '2020-11_Nov-Dec', 'figures', 'fraction-mapped_fit-db-1.png'))
```
]

???
- _De novo_ and open-reference perform equally well.
- Closed reference performs slightly worse.
- RDP doesn't perform well.
- Prior to switching to silva v132 & updating mothur, closed-reference artificially outperformed open-reference.

---

# Reference clustering against a database

.pull-left[
```{r runtime_fit-db}
include_graphics(here('exploratory', '2020-11_Nov-Dec', 'figures', 
                      'runtime_fit-db-1.png'))
```
]

--

.pull-right[
```{r memory_fit-db}
include_graphics(here('exploratory', '2020-11_Nov-Dec', 'figures', 
                      'memory_fit-db-1.png'))
```
]

---

## Summarizing mothur results

Whenever there's virtually no variation in the metric.

```{r sum_all_mothur}
include_graphics(here('exploratory', '2020-11_Nov-Dec', 'figures', 
                      'mcc_all-opti-1.png'))
```

---

## mothur vs. vsearch

```{r vsearch}
include_graphics(here('exploratory', '2020-11_Nov-Dec', 'figures', 
                      'metrics_all-human-1.png'))
```

???
- Just showing the human dataset here because results patterns were basically the same across all datasets.
- Only used greengenes for reference because that's what
QIIME does.

---

# Splitting datasets into reference & query sets

.pull-left[
Before reverting to Silva v132

```{r frac-map-old}
include_graphics(here('exploratory', '2020-05_May-Oct', 'figures', 
                      'fraction_mapped_complement-2.png'))
```
]
--
.pull-right[
After reverting to Silva v132:

```{r frac-map-new}
include_graphics(here('exploratory', 
                      '2020-11_Nov-Dec', 
                      'figures', 
                      'debug-fraction-mapped_fit-split-1.png'))
```
]

???
- 1 seed out of every parameter set has the same pattern as the old results
- I thought maybe my snakemake workflow somehow kept one result file from before reverting to silva. Like maybe it accidentally didn't overwrite with the new results?
  - But I checked the timestamps of the results files and they were updated after re-doing with silva v132.
---

## Sanity check: open-ref should have fraction mapped = 1

.pull-left[
```{r}
include_graphics(here('exploratory', 
                      '2020-11_Nov-Dec', 
                      'figures',
                      'debug-fraction-mapped-open_fit-split-1.png'))
```
]

???
- Now that is really weird. For weight=simple, the fraction mapped is equal to the sample frac most of the time. So the fraction mapped is incorrect most of the time.
- Next I thought; maybe there's a bug in mothur where the open-ref output files have one less column than the closed-ref.
  - I brought this up to Sarah W before and she said she fixed it. Maybe I forgot to recompile mothur?
  - Bingo! I hadn't pulled the latest changes. So I recompiled and re-ran with one parameter set & one seed, and it worked! Problem solved! Well, not quite. When re-running with more than one seed, the problem persisted.
---

## is there a pattern?

```{r}
dat_100 <- read_tsv(here('exploratory',
                               '2020-11_Nov-Dec',
                               'debug-split_100-seeds.tsv'))
dat_100 %>% filter(method == 'open') %>% 
  mutate(is_correct = fraction_mapped == 1) %>% 
  select(dataset, seed, is_correct, fraction_mapped,
         sample_frac, ref_frac, ref_weight) %>% 
  filter(is_correct, dataset == 'soil')
```

???
- is there a bug in my code to calculate the fraction mapped? let's check results from fitting to databases

---

## Sanity check: results from fitting to databases

.pull-left[
```{r}
include_graphics(here('exploratory', 
                      '2020-11_Nov-Dec', 
                      'figures',
                      'debug-fraction-mapped-open_fit-split-1.png'))
```
]
.pull-right[
```{r}
include_graphics(here('exploratory', 
                      '2020-11_Nov-Dec', 
                      'figures',
                      'fraction-mapped-open_fit-db-1.png'))
```
]
???
- so there isn't a bug in my fraction mapped code and there isn't a bug in mothur, otherwise these would be wrong too.
- is there a bug in my code to split the datasets into reference & query fractions?
  - I had written assertion statements throughout that code to make sure the query + reference fraction were correct.
  - Double-checked the resulting accnos files anyway just to be safe. They were all correct.
---

## WHAT?!

```{r}
split_2 <- read_tsv(here('exploratory',
                          '2020-11_Nov-Dec',
                          'debug-split_2-seeds.tsv')) %>%
  filter(method == 'open') %>% 
  mutate(is_correct = fraction_mapped == 1) %>% 
  select(dataset, seed, method, fraction_mapped, is_correct, 
         sample_frac, ref_frac, ref_weight)

split_2 %>% 
  filter(dataset == 'soil', ref_weight == 'simple') %>% 
  select(sample_frac, ref_frac, seed, is_correct) %>% 
  pivot_wider(names_from = seed, 
              values_from = is_correct,
              names_glue = 'seed_{seed}_is_correct')
```

Re-ran with only 2 seeds.

---

```{bash, eval = FALSE}
wget https://raw.githubusercontent.com/SchlossLab/OptiFitAnalysis/c900c5fa5f3081b73536a4e568f66817a66b3d0c/figures/rulegraph_fit_split.png?token=AEHR6TPEQM4HG6GCTVOUSTK74DIXK -O exploratory/2020-11_Nov-Dec/figures/rulegraph1.png
```
.pull-left[
Workflow

```{r rulegraph_old, out.width='60%'}
include_graphics(here('exploratory', 
                      '2020-11_Nov-Dec', 
                      'figures',
                      'rulegraph1.png'))
```
]
--
.pull-right[
So far:
- Code shared with  workflow that fits to databases must be correct.
- Code that didn't change between previous and most recent results must be correct.
  - What code did I change between then and now? Did I change anything?
]
???
- Browsed my commit history on github and realized that I had done a little refactoring to clean up the workflow a bit. 
- I also added the rule pick_best_cluster_seed to check off a TODO in the analysis plan. I forgot about that, thought that had been there much longer.
- So I removed that rule and tried re-running the workflow with one parameter set and 2 seeds to check whether that was the problem.

---

## Fixed it!
.pull-left[
```{r rulegraph_new, out.width='60%'}
include_graphics(here('figures',
                      'rulegraph_fit_split.png'))
```
]
.pull-right[
```{r}
include_graphics(here('exploratory', 
                      '2020-11_Nov-Dec', 
                      'figures',
                      'fraction-mapped-grid_fit-split-1.png'))
```
]

---

## What happened here?

.pull-left[
- With `pick_best_cluster_seed`, the OptiClust results from the best seed were used for downstream OptiFit runs with _all seeds_.
  - Same seed variable used to create the data splits.
- When seeds mismatched, query sets were being fit to reference OTUs from different data splits.
- Results were correct whenever the OptiFit seed corresponded to the best OptiClust seed, and incorrect in all other cases.
]

---

## Conclusions

- Don't change more than one thing at a time, even if you think it's "just refactoring".
- Start debugging from the easiest step to fix.
- Have config files with different parameter sets to make debugging quick & easy.
  - default: all datasets & parameters & 100 seeds.
  - debug: one dataset & only one of each parameter choice & 2 seeds.
- Make many small commits to make it easy to figure out where you introduced a bug.
- Test your assumptions
  - Use `assertthat` pkg in R.