### 2020-12-15 exploratory/2020-11_Nov-Dec/plots.Rmd
Plot de novo stats as hline

---
date: 'Nov. - Dec. 2020'
output: 
  github_document:
    html_preview: false
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      fig.path = "figures/"
                      )
```
```{r lib}
set.seed(2018)
library(cowplot)
library(ggtext)
library(glue)
library(here)
library(knitr)
library(tidyverse)

theme_set(theme_classic())
color_palette <- RColorBrewer::brewer.pal(4, "Dark2")
dataset_colors <- c(
  human = color_palette[[3]],
  marine = color_palette[[1]],
  mouse = color_palette[[4]],
  soil = color_palette[[2]]
)
tri_colors <-  c("darkorange","darkorchid","cyan4") # https://allisonhorst.github.io/palmerpenguins/articles/intro.html
mutate_perf <- function(dat) {
  dat %>% 
    mutate(mem_mb = max_rss,
           mem_gb = mem_mb / 1024) %>% 
    rename(sec = s,
           num_otus = sobs)
}
```

## _de novo_ clustering

```{r load_opticlust}
opticlust <- read_tsv(here('subworkflows/1_prep_samples/results/opticlust_results.tsv')) %>% 
  full_join(read_tsv(here('subworkflows/1_prep_samples/results/dataset_sizes.tsv'))) %>% 
  mutate_perf()
sum_opticlust <- opticlust %>% 
  group_by(dataset) %>% 
  summarize(mcc_median = median(mcc),
            sec_median = median(sec),
            mem_gb_median = median(mem_gb))#,
            #median_shannon = median(shannon))
```
```{r mcc_opticlust}
opticlust %>% 
  ggplot(aes(dataset, mcc)) +
  geom_jitter(alpha = 0.5) +
  ylim(0.5, 1)
```
```{r runtime_opticlust}
opticlust %>% 
  ggplot(aes(num_seqs, sec, color = dataset)) +
  geom_point(alpha = 0.5)
```
## fit to reference databases

```{r load_fit-dbs}
ref_sizes <- read_tsv(here('subworkflows/2_fit_reference_db/results/ref_sizes.tsv'))
optifit_dbs <- read_tsv(here('subworkflows/2_fit_reference_db/results/optifit_dbs_results.tsv')) %>% 
  mutate_perf()
head(optifit_dbs)
```

```{r mcc_fit-db}
optifit_dbs %>% 
  ggplot(aes(x=method, y=mcc, color=ref)) +
  geom_hline(aes(yintercept = mcc_median), sum_opticlust) +
  geom_boxplot() + 
  facet_wrap('dataset') +
  scale_color_manual(values=tri_colors) +
  ylim(0, 1) +
  labs(title='OTU Quality: OptiFit with reference databases',
       caption='Black line: _de novo_ clustering') +
  theme(plot.caption = element_markdown())
```

```{r fraction-mapped_fit-db}
optifit_dbs %>% filter(method == 'closed') %>% 
  ggplot(aes(x=dataset, y=fraction_mapped, color=ref)) +
  geom_boxplot() + 
  facet_wrap('method') +
  ylim(0, 1) +
  scale_color_manual(values=tri_colors) +
  labs(title="Sequences mapped during closed-reference OptiFit") +
  theme(plot.caption = element_markdown())
```

```{r runtime_fit-db}
optifit_dbs %>% 
  ggplot(aes(x=method, y=sec, color=ref)) +
  geom_boxplot(alpha = 0.5) +
  geom_hline(aes(yintercept = sec_median), sum_opticlust) +
  facet_wrap('dataset', scales = 'free') +
  scale_color_manual(values=tri_colors) +
  labs(title='OptiFit runtime with reference databases',
       caption='Black line: _de novo_ clustering') +
  theme(plot.caption = element_markdown())
```

```{r memory_fit-db}
optifit_dbs %>% 
  ggplot(aes(x=method, y=mem_gb, color=ref)) +
  geom_boxplot(alpha = 0.5) +
  geom_hline(aes(yintercept = mem_gb_median), sum_opticlust) +
  facet_wrap('dataset', scales = 'free') +
  scale_color_manual(values=tri_colors) +
  labs(title='OptiFit runtime with reference databases',
       caption='Black line: _de novo_ clustering') +
  theme(plot.caption = element_markdown())
```

```{r diversity_fit-db, eval=FALSE}
optifit_dbs %>% 
  ggplot(aes(x=method, y=shannon, color=ref)) +
  geom_jitter(alpha = 0.5) +
  geom_hline(aes(yintercept = median_shannon), sum_opticlust) +
  facet_wrap('dataset', scales = 'free') +
  scale_color_manual(values=tri_colors) +
  labs(title='OptiFit runtime with reference databases',
       caption='Black line: _de novo_ clustering') +
  theme(plot.caption = element_markdown())
```

## fit split datasets

```{r load_fit-split}
optifit_split <- read_tsv(here('subworkflows/3_fit_sample_split/results/optifit_split_results.tsv')) %>% 
  mutate_perf()
```

```{r mcc_fit-split}
optifit_split %>% 
  ggplot(aes(x = ref_frac, y = mcc, color = ref_weight)) +
  geom_jitter(size = 1, alpha = 0.3, width = 0.01) +
  stat_summary(fun = mean, geom = 'crossbar', alpha = 0.5) +
  geom_hline(aes(yintercept = mcc_median), sum_opticlust) +
  facet_grid(dataset ~ method) +
  ylim(0, 1) +
  labs(title='OTU Quality: OptiFit with split datasets',
       x='reference fraction',
       caption='Black line: _de novo_ clustering on the whole dataset') +
  theme(plot.caption = element_markdown())
```

```{r runtime_fit-split}
optifit_split %>% 
  ggplot(aes(x = ref_frac, y = sec, color = ref_weight)) +
  geom_jitter(size = 1, alpha = 0.3, width = 0.01) +
  stat_summary(fun = mean, geom = 'crossbar', alpha = 0.5) +
  geom_hline(aes(yintercept = sec_median), sum_opticlust) +
  facet_grid(dataset ~ method, scales = 'free') +
  labs(title='Runtime: OptiFit with split datasets',
       x='reference fraction',
       caption='Black line: _de novo_ clustering on the whole dataset') +
  theme(plot.caption = element_markdown())
```

### fraction mapped

```{r fraction-mapped_fit-split}
optifit_split %>% 
  filter(method == 'closed', tool == 'mothur', is.na(ref)) %>% 
  #group_by(dataset, ref_weight, ref_frac) %>% 
  ggplot(aes(x=ref_frac, y=fraction_mapped, color=ref_weight)) +
  geom_point(alpha = 0.5, size=1) +
  facet_wrap("dataset", nrow=1) +
  ylim(0, 1) +
  labs(title="Sequences mapped during closed-reference OptiFit",
       x='reference fraction',
       y='fraction mapped')
```

It looks like one result per parameter set is left over from before I re-ran 
everything after reverting to the older silva version.
[Here's what the plot looked like with the newest (buggy) silva](https://github.com/SchlossLab/OptiFitAnalysis/blob/master/exploratory/2020-05/sub3_fit_all-seqs.md#fraction-of-sequences-that-map-to-the-reference).
Need to go back and figure out why those files didn't get removed/overwritten
by Snakemake when I re-ran everything.

Let's try a sanity check with open-reference -- should all be 1.

```{r fraction-mapped-grid_fit-split}
optifit_split %>% 
  filter(tool == 'mothur', is.na(ref)) %>% 
  filter(dataset == 'soil') %>% 
  ggplot(aes(x=ref_frac, y=fraction_mapped, color=ref_weight)) +
  geom_point(alpha = 0.5, size=1) +
  facet_grid(dataset ~ method) +
  ylim(0, 1) +
  labs(title="Sequences mapped during OptiFit",
       x='reference fraction',
       y='fraction mapped',
       caption=glue("Ran with {optifit_split %>% pull(seed) %>% unique() %>% length()} seeds"))
```

Did mothur's list file format change in the latest dev version? The fraction
mapped values are the same as the reference fraction when the ref weight is
simple.

### is there a pattern?

```{r debug_frac_map}
debug_dat <- optifit_split %>% 
  filter(method == 'open') %>% 
  mutate(is_correct = fraction_mapped == 1) %>% 
  select(dataset, seed, method, fraction_mapped, is_correct, 
         sample_frac, ref_frac, ref_weight) %>%
  group_by(dataset, seed, ref_frac, ref_weight, is_correct)

for (s in 1:2) {
  debug_dat %>% 
    filter(seed == s, ref_weight == 'simple', dataset == 'soil') %>% 
    pull(is_correct) %>% 
    print()
}

```

Pattern of correctness among ref fracs is flipped between seed 1 and 2

### diversity

```{r diversity_fit-split, eval = FALSE}
optifit_split %>% 
  ggplot(aes(x = ref_frac, y = shannon, color = ref_weight)) +
  geom_jitter(size = 1, alpha = 0.3, width = 0.01) +
  stat_summary(fun = mean, geom = 'crossbar', alpha = 0.5) +
  geom_hline(aes(yintercept = median_shannon), sum_opticlust) +
  facet_grid(dataset ~ method, scales = 'free') +
  labs(title='Diversity: OptiFit with split datasets',
       x='reference fraction',
       caption='Black line: _de novo_ clustering on the whole dataset') +
  theme(plot.caption = element_markdown())
```

## vsearch

For reference-based clustering, datasets were fit to the greengenes database.

```{r load_vsearch}
vsearch <- read_tsv(here('subworkflows/4_vsearch/results/vsearch_results.tsv')) %>% 
  mutate_perf()
vsearch_join <- vsearch %>% 
  full_join(opticlust) %>% 
  full_join(optifit_dbs %>% filter(ref == 'gg'))
```

```{r mcc_vsearch}
vsearch_join %>% 
  ggplot(aes(x = method, y = mcc, color = tool)) +
  geom_boxplot() +
  facet_wrap('dataset') +
  ylim(0, 1) + 
  labs(x = '')
```

```{r sec_vsearch}
vsearch_join %>% 
  ggplot(aes(x = method, y = sec, color = tool)) +
  geom_boxplot() +
  facet_wrap('dataset') +
  labs(x = '')
```

```{r diversity_vsearch}
vsearch_join %>% 
  ggplot(aes(x = method, y = shannon, color = tool)) +
  geom_boxplot() +
  facet_wrap('dataset') +
  labs(x = '')
```

```{r fraction-mapped_vsearch}
vsearch_join %>% #filter(method == 'closed') %>% 
  ggplot(aes(x = dataset, y = fraction_mapped, color = tool)) +
  geom_boxplot() + 
  ylim(0, 1) +
  facet_wrap('method') +
  labs(title="Sequences mapped during closed-reference OptiFit") +
  theme(plot.caption = element_markdown())
```

The fraction mapped for _de novo_ open-reference is shown here as a sanity check.
Final plots will only include closed-reference.

## compare all mothur clustering strategies

```{r join_all-opti}

plot_denovo_hline <- function(yint, dat = sum_opticlust) {
  list(geom_hline(data = dat, aes(yintercept = {{yint}})),
       labs(caption = "Black line: _de novo_ clustering0"),
       theme(plot.caption = element_markdown())
  )
}

optifit_all <- list(optifit_dbs %>% 
                   mutate(strategy = glue('database_{ref}')),
                 optifit_split %>% 
                   mutate(strategy = 'self-split')) %>% 
  reduce(full_join)

sum_optifit <- optifit_all %>% 
  group_by(dataset, strategy, method) %>% 
  summarize(n = n(),
            mcc_median = median(mcc),  # TODO: tidy way to avoid this repetitiveness?
            sec_median = median(sec),
            mem_gb_median = median(mem_gb),
            frac_map_median = median(fraction_mapped))
head(sum_optifit)
```

```{r mcc_all-opti, fig.width=10, fig.height=3.5}
sum_optifit %>% 
  ggplot(aes(strategy, mcc_median, color = method)) +
  geom_point() +
  facet_wrap(dataset ~ ., nrow=1) +
  scale_color_manual(values = tri_colors) +
  ylim(0, 1) +
  #theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + 
  coord_flip() +
  labs(title = 'OTU Quality', x = '', y = 'Median MCC') + 
  plot_denovo_hline(yint = mcc_median)

```
```{r runtime_all-opti}
sum_optifit %>% 
  ggplot(aes(strategy, sec_median, color = method)) +
  geom_point() +
  facet_wrap(dataset ~ ., scales = 'free_y') +
  scale_color_manual(values = tri_colors) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  plot_denovo_hline(sec_median)
```
```{r memory_all-opti}
sum_optifit %>% 
  ggplot(aes(strategy, mem_gb_median, color = method)) +
  geom_point() +
  facet_wrap(dataset ~ ., scales = 'free_y') +
  scale_color_manual(values = tri_colors) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  plot_denovo_hline(mem_gb_median)
```

```{r fraction-mapped_all-opti}
sum_optifit %>% 
  ggplot(aes(strategy, frac_map_median, color = method)) +
  geom_point() +
  facet_wrap(dataset ~ .) +
  scale_color_manual(values = tri_colors) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

## compare mothur & vsearch with just the human dataset

```{r join_tools_human}
all_human <- list(optifit_all %>% 
                    filter(dataset == 'human', 
                           ref == 'gg' | method == 'de_novo'),
                  opticlust %>% filter(dataset == 'human'),
                  vsearch %>% filter(dataset == 'human')
                 ) %>% 
  reduce(full_join) %>% 
  mutate(strategy = case_when(
    method == 'de_novo' ~ method,
    TRUE ~ as.character(glue('{method}_ref_{ref}'))))

sum_all_human <- all_human %>% 
  group_by(tool, strategy, method) %>% 
  summarize(n = n(),
            mcc_median = median(mcc),  # TODO: tidy way to avoid this repetitiveness?
            sec_median = median(sec),
            mem_gb_median = median(mem_gb),
            frac_map_median = median(fraction_mapped))
head(sum_all_human)
```

```{r mcc_all-human}
sum_all_human %>% 
  ggplot(aes(strategy, mcc_median, color = tool)) +
  geom_point() +
  #facet_wrap(tool ~ .) +
  ylim(0, 1) +
  scale_color_manual(values = tri_colors) +
  #theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + 
  coord_flip()
```

```{r sum_all-human}
sum_all_human %>% 
  pivot_longer(c(mcc_median, sec_median, mem_gb_median, frac_map_median), names_to = 'metric') %>% 
  ggplot(aes(strategy, value, color = tool)) +
  geom_point() +
  facet_wrap('metric', scales = 'free')

```

```{r metrics_all-human, fig.width=10, fig.height=3.5}
all_human %>% 
  pivot_longer(c(mcc, sec, mem_gb, fraction_mapped), names_to = 'metric') %>% 
  ggplot(aes(strategy, value, color = tool)) +
  geom_boxplot() +
  facet_wrap('metric', scales = 'free_y', nrow=1) +
  labs(x='', y='',  
       caption = 'Only the human dataset is shown here. The patterns are similar with other datasets.') +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
```