### 2021-03-03 docs/slides/2021-03-11_bistro.Rmd
Create file for BISTRO pres

---
title: "OptiFit project update"
subtitle: "+ a case study in debugging"
author: "Kelly Sovacool"
date: "2021-03-11 (updated: `r Sys.Date()`)"
output:
  xaringan::moon_reader:
    css: libs/xaringan-themer.css
    lib_dir: libs
    nature:
      ratio: '16:9'
      highlightStyle: github
      highlightLines: true
      #titleSlideClass: ['center', 'middle', 'title-slide']
    chakra: libs/remark-latest.min.js
    self_contained: true
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE
                  )
# prev slides:
# - Oct 2020 research talk https://docs.google.com/presentation/d/1GbbCYmlF4jkrO5BGyCTgtFBOL4pfzPDIQLMZBND0Hzg/edit#slide=id.p
# - Dec 2020 research talk https://github.com/SchlossLab/OptiFitAnalysis/blob/master/docs/slides/2020-12-14_lab-mtg.Rmd
# - fall 2019 JC https://github.com/kelly-sovacool/bioinf603-fall2019

# TODO: use http urls instead of here filepaths for figures
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_light(
  header_font_google = google_font("Helvetica"),
  text_font_google = google_font("Helvetica")
)
# style_mono_light(base_color = "#23395b")
# style_mono_accent(
#   base_color = "#1c5253",
#   header_font_google = google_font("Josefin Sans"),
#   text_font_google   = google_font("Montserrat", "300", "300i"),
#   code_font_google   = google_font("Fira Mono")
# )
```

```{r chakra, eval = FALSE}
# run once to download remark-lastest
xaringan::summon_remark(to = 'docs/slides/libs')
```

```{r libs}
library(here)
library(knitr)
library(tidyverse)
xaringanExtra::use_share_again()
```

## OTU Clustering methods

- _De novo_ (OptiClust)
  - Cluster sequences based on their similarity.
  - OTU's aren't stable; you'll get different OTU assignments with different random seeds.
- Reference-based (OptiFit)
  - First cluster reference sequences _de novo_, then fit query sequences to those OTUs.

.footnote[
  [Westcott & Schloss (2015) _PeerJ_ ](https://doi.org/10.7717%2Fpeerj.1487);
  [Schloss (2016) _mSystems_ ](https://doi.org/10.1128%2FmSystems.00027-16);
  [Westcott & Schloss (2017) _mSphere_](10.1128/mSphereDirect.00073-17)
]

--

### Reference-based strategies

- Modes
  - Open - sequences that can't be fit to the reference are then clustered _de novo_ to create additional OTUs.
  - Closed - sequences that can't be fit to the reference are thrown out.
- Reference choice
  - External database (Silva, greengenes, RDP).
  - Split the dataset into a reference & query set.

---

![](../../exploratory/2020-11_Nov-Dec/figures/mcc_fit-db-1.png)