### 2021-03-24 exploratory/2021/test_criteria.Rmd
Plot mcc & fraction mapped from criteria test

---
date: '`r Sys.Date()`'
output: 
  github_document:
    html_preview: false
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      fig.path = "figures/"
                      )
```

```{r lib}
set.seed(2018)
library(cowplot)
library(ggtext)
library(glue)
library(here)
library(knitr)
library(tidyverse)

theme_set(theme_bw())

mutate_perf <- function(dat) {
  dat %>% 
    mutate(mem_mb = max_rss,
           mem_gb = mem_mb / 1024) %>% 
    rename(sec = s)
}
```

```{r read}
dat <- read_tsv(here('subworkflows', '3_fit_sample_split', 'results', 
                     'optifit_crit_check.tsv')) %>% 
  mutate_perf() %>% 
  select(dataset, printref, criteria, mcc, fraction_mapped, sec, mem_gb)
```

## OTU Quality

```{r crit_mcc}
dat %>% 
  ggplot(aes(x = criteria, y = mcc, color = printref)) +
  geom_point(alpha = 0.5) +
  ylim(0, 1)
```
## Fraction mapped

```{r crit_map}
dat %>% 
  ggplot(aes(x = criteria, y = fraction_mapped, color = printref)) +
  geom_point(alpha = 0.5) +
  ylim(0, 1)
```