### 2021-10-21 exploratory/2021/seq_stats.Rmd
Plot similarity & abundance of seqs

---
date: '`r Sys.Date()`'
output: 
  github_document:
    html_preview: false
---

# Similarity & Abundance of Sequences

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      fig.path = "figures/",
                      dpi = 300
                      )
```
```{r deps, warning=FALSE, message=FALSE}
set.seed(20211021)
library(glue)
library(here)
library(tidyverse)
theme_set(theme_bw())
```

```{r dat}
seq_stats <- c("human","marine","mouse","soil") %>% map_dfr(function(x) {
  read_tsv(glue('subworkflows/1_prep_samples/data/{x}/seq_stats.tsv')) %>% 
    mutate(dataset = x)
}) %>% pivot_longer(c(similarity, abundance), names_to = 'statistic')
```

```{r hist_sim_abun}
seq_stats %>% 
  ggplot(aes(x=value, fill=statistic)) +
  geom_histogram() +
  facet_grid(dataset ~ statistic, scales='free')
```

## log10

```{r hist_sim_abun_log10}
seq_stats %>% 
  ggplot(aes(x=value, fill=statistic)) +
  geom_histogram() +
  scale_x_log10() +
  facet_wrap("dataset")
```

## abundance vs similarity

```{r point_abun_sim}
seq_stats %>% 
  pivot_wider(names_from = statistic) %>% 
  ggplot(aes(x = abundance, y = similarity, color = dataset)) +
  geom_point(alpha=0.7) +
  scale_x_log10() +
  facet_wrap('dataset')
```