### 2021-11-09 paper/head.tex

\fancyhead[L]{\small{OptiFit: improved reference-based clustering}}
\fancyhead[R]{\small{Sovacool \textit{et al.}}}

\begin{center}
\vspace{25mm}

Kelly L. Sovacool${^1}$, Sarah L. Westcott${^2}$, M. Brodie Mumphrey${^1}$, Gabrielle A. Dotson${^1}$, Patrick D. Schloss$^{2,3,\dagger}$

\vspace{20mm}

$1$ Department of Computational Medicine and Bioinformatics, University of Michigan

$2$ Department of Microbiology and Immunology, University of Michigan

$3$ Center for Computational Medicine and Bioinformatics, University of Michigan

\vspace{30mm}

$\dagger$ To whom correspondence should be addressed: pschloss@umich.edu

\end{center}

\newpage
\linenumbers